The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-02 carbon_paper/tex/main.tex
more developed intro

% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols
\graphicspath{{../figures/}}

@@

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\defcitealias{cristallo+11}{C11}
\defcitealias{cristallo+15}{C15}
\defcitealias{ventura+13}{V13}

\newcommand{\cristallo}{\citetalias{cristallo+11}+\citetalias{cristallo+15}}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
% \newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\VICE}{\texttt{VICE}}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Constraining carbon yields]{The galactic chemical evolution of carbon: Constraining stellar nucleosynthesis}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et al.]{
Daniel A. Boyea,$^{1}$\
et al.
% James W. Johnson,$^{1}$
% David Weinberg
% Jack Roberts
% Jennifer Johnson
% ?
\\
% List of institutions
$^{1}$Department of Astronomy, The Ohio State University\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

% Abstract of the paper
\begin{abstract}
Stellar evolution models provide highly uncertain predictions for elemental yields due to our limited understanding of stellar physics. In this paper, we aim to estimate the nucleosynthetic yields of carbon using galactic chemical evolution models.

We find that the AGB stars make up a fraction $0.20_{-20}^{+80}$ of total carbon production at late times. 
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
Need to find some...
\end{keywords}

@@

A fundamental question of Galactic Chemical Evolution (GCE) is understanding the physical processes which make each element. Stellar evolution models can be used to predict how much of each element stars with a variety of physical properties produce. However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, opacity, and mass loss (CITE, e.g. stellar physics reviews).

Carbon and nitrogen are the only lighter elements produced significantly by low mass stars, so*****, reasons WHY this study.

Asymptotic Giant Branch (AGB) stars are known to be importaint (cite) from stellar model predictions, previous GCE, and evidence from abundance patterns. (are there detections in PN?)

CCSNe are also important (.... SN reminant measurments)

We also know that there is evidence for non-monatonic carbon yield patterns. 
pattern where carbon is more readily produced by both high metallicity and very low metallicity stars (CEMPS, early GCE results), but our picture.

Additionally, stellar models are computationally expensive, even in one dimension, so while there are grids of predictions, these are not always fine enough to make accurate chemical evolution models.

Previous models (CIte). We use \VICE (CITE) to create multi-zone models of GCE enabling us to account for the effects of migration on our predicted yields.

\citet{prantzos+18}
\citep{WAF17}
\citep{james+21}
\citep{fiorenzo+21}

\citep[e.g.][]{james+22}

Carbon abundances are strongly affected by CNO cycle processing in stars. The CNO cycle is one of two series of nuclear reactions( $^{12}$C(p, $\gamma$)$^{13}$N($\beta^+ \nu_e$)$^{14}$N(p, $\gamma$)$^{15}$O($\beta^+\nu_e$)$^{15}$N(p,$\alpha$)$^{12}$C

Because the $^{14}$N(p, $\alpha$) proton capture is the slowest component of the CNO cycle ****, the CNO cycle, to first order, converts all $^{12}$C into $^{14}N$.

One challenge of measuring abundances of carbon in star is that when a star becomes a RGB star, material from the CNO-processed core. Since this core-processed material is now carbon depleted and nitrogen enhanced, measurments of this star's atmosphere will no longer reflect the birth abundances of the star.

In Figure \ref{fig:sum} we present a sample of APOGEE subgiants. These subgiants were chosen based on their location in the HR diagram based on criteria in \citet{jack_subgiant} replicated in Appendix \ref{sec:jack}. Because subgiants have not undergone FDU, their surface abundances are an accurate reflection of the birth abundances (CITATION).

\begin{figure*}
    \centering
    \includegraphics[]{subgiants_mgh.pdf}
    \includegraphics[]{subgiants_mgfe.pdf}
    \caption{Chemical abundance relationships of our sample of APOGEE subgiants (see Jack et al. 2023). Left panel shows }
    \label{fig:sum}
\end{figure*}
\section{Nucleosynthesis}

Galactic chemical evolution are fundamentally intertwined with our understanding of stellar nucleosynthesis. In this paper, we are concerned with three major nucleosynthetic pathways: AGB stars, CCSNe, and Supernovae type Ia. While only AGB stars and CCSNe are producers of carbon, SNeIa Fe can be used to compare with the AGB delayed carbon component.

\VICE allows us to explore the nucleosynthesis of various models, as it includes AGB tables for C11, et.c

We discuss carbon nucleosythesis in more detail in the next two sections. We assume that the only dominant sources of carbon are from AGB stars and CCSNe/massive star winds. ***JUSTIFY***. Below  is a brief overview of our  nucleosythetic assumptions of other elements.

Let $y_{\rm X}^{\rm proc}(M, Z)$ be the fractional yield of element X by a process proc for stars of a mass $M$ (in solar masses) and at metallicity $Z$ defined as: 
\begin{equation}
    y_{\rm X} = \frac{M_{\rm X,\ ej} - M_{\rm X,\ birth}}{M}   
\end{equation}

Where $M_{\rm X,\ ej}$ and $M_{\rm X,\ birth}$ are the mass of element X ejected and initially present in the star. 
We define the IMF-integrated yield of X with: 
\begin{equation}
Y_{\rm X}^{\rm proc}(Z) = \int_{M_{\rm min}}^{M_{\rm max}} y_{\rm X}^{\rm proc}(M, Z)\ \Phi(M)  \ dM
\end{equation}
Where $\Phi(M) = \frac{dN}{dM}/ \int_{M_{\rm min}}^{M_{\rm max}} \frac{dN}{dM}\ dM$ is the normalized IMF, $M_{\rm min}$ and $M_{\rm max}$ are the minimum and maximum mass of stars, which we take to be $0.08 M_{\sun}$ and $100 M_{\sun}$ respectively; and $\frac{dN}{dM}$ is the initial mass function, the frequency of stars born between $M$ and $M+dM$.

There are two $\alpha$-elements we care about: Mg and O. Stellar observations of magnesium are more reliable but oxygen abundances are easier to measure in HII regions. We assume that both Mg and O are produced entirely in CCSNe with constant, metallicity-independent yields. Thus $\text{[Mg/H]} = \text{[O/H]} = [\alpha/\text{H}]$ in our models

\begin{itemize}
    \item observational validation
\end{itemize}
For carbon, the dominant producer is from massive stars which eject carbon fused in their cores during CCSNe.

Taken from J21 and J22, we make the following yield assumptions:
\begin{itemize}
    \item $Y_\text{O}^\text{CC} = 0.015$
    \item $Y_\text{O}^\text{Ia} = 0$
    \item $Y_\text{Fe}^\text{Ia} = 0.00214?$
    \item $Y_\text{Fe}^\text{CC} = 0.0012$
    \item $Y_\text{N}^\text{CC} = 0.00036$
    \item $y_\text{N}^\text{AGB}(M, Z) = 9\times 10^{-4} M \left(\frac{Z}{Z_\odot}\right)$
\end{itemize}
Also following J21/J22, we take the SNeIa Delay Time distribution to be a $t^{-1.1}$ law based on
the observations by \citet{maoz+12}.

We leave consistent throughout the paper. While different CCSNe models predict different absolute values for $y_\text{O}^\text{CC}$, they also predict that the yield is metallicity-independent within the uncertainties of the model.

\subsection{Asymptotic giant branch carbon}

An AGB star is a low mass ($\lesssim 8 M_{\sun}$) star during the final phases of evolution. The star has already completed Hydrogen core burning, Hydrogen shell burning during the red giant branch phase, and Helium core burning. So, the structure consists of an inert core of carbon and oxygen surrounded by a Helium burning shell and a Hydrogen burning shell. The double shell burning leads to thermal instabilities which drive off the outer layers of the star in a series of violent pulsations, ejecting the envelope and leaving behind a C/O white dwarf.

In an AGB star, TDU and HBB are two major processes involved in carbon production and destruction. TDU occurs during thermal pulsations, where the convective layer of the star passes into the inner layers with He-burning products which mixes in C and O into the envelope. In contrast, HBB may happen if the base of the convective envelope reaches about 50MK, initiating the CNO cycle, converting most C12 into N14.

The differences in yields between AGB models depend on the mass loss rate, treatment of convection, and nuclear reaction rates.

In this work, we explore four different sets of AGB star yields representative of the variety of options which also provide reasonably well sampled grids in metallicity and mass necessary for chemical evolution studies.
\begin{itemize}
    \item \citetalias{cristallo+11}
    \item \cristallo
    \item \cite[C11][Hi]{cristallo+11}: \citet{cristallo+11}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16}
\end{itemize}
Unless otherwise noted, C11 is the model we choose as our fiducial AGB yield.

Figure \ref{fig:y_agb} compares the net fractional AGB carbon yield, $m_\text{C}^\text{AGB}(M, Z)$ for each AGB model we consider. Note that the yields may be negative when the birth abundance is higher than the carbon abundance of the material ejected back to the ISM.

Most models agree on the qualitative shape of the net fractional AGB carbon yield. The stars with the highest fractional carbon yields are between 2 to 4 solar masses, depending on the metallicity and model. Both lower and higher mass stars produce less carbon, and high mass stars may even destroy carbon at high metallicity due to highly efficient HBB. And all models agree that the carbon yield generally decreases with metallicity.

Metallicity dependence:
V13 is the only model that stands out, showing a non-monotonic metallicity dependence, however this effect is only for models with [M/H]$\leq -1$. Otherwise the major differences between the models are the precise amount of carbon and the slope of the metallicity dependence. For example, the three models C11, K10, and K16 predict $y_\text{C}^\text{AGB}$ to be between 0.006 and 0.008 at solar metallicity, but C11 has a much shallower metallicity dependence than the K10 and K16 models. And at solar, V13 predicts a value at around 0.004 so the models are all within about a factor of 2.

Another difference is the speed at which equilibrium is reached. K10 and K16 weight carbon production more heavily towards massive stars resulting in a faster time to equilibrium, whereas the C11 and V13 models predict a slightly longer timescale with order 1Gyr, but little to no carbon is produced more than 2Gyr after a star formation event. This is in contrast to iron which is still produced by SNeIa up to 10Gyr after a star formation event.

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}\\

    \caption{The net fraction carbon yield ($M_\text{C}$, the amount of carbon produced by the star divided by $M$) plotted as a function of mass, $M$. Lighter colors represent higher values of metalicity ([M/H]). Each panel represents a different AGB model. K10 reports stars at MoverH -3, -2, etc. I can reproduce C11 and K10 but not other V13 and KL16 :(((}

    \label{fig:y_agb}
\end{figure*}

\begin{figure*}
    \centering

    \includegraphics[scale=1]{y_agb_vs_z-1.pdf}
    \includegraphics[scale=1]{y_agb_vs_t-1.pdf}

    \caption{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

\subsection{CCSNe carbon}

While there are many stellar models providing predictions of CCSNe yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

We summarise many of the predictions of stellar models in Figure \ref{fig:y_cc}. Pink squares
represent \citep{NKT13}, etc. etc.
\citep{LC18}
\citep{sukhbold+16}
\citep{WW95}

>>>>

% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols

\usepackage{isotope}

\usepackage{hyperref}

\graphicspath{{../figures/}}

@@

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

\newcommand{\cc}{CCSNe}
\newcommand{\agb}{AGB}
\newcommand{\ia}{SNe Ia}
\newcommand{\sfh}{SFH}
\newcommand{\dla}{DLA}
\newcommand{\ssp}{SSP}
\newcommand{\imf}{IMF}
\newcommand{\gce}{GCE}
\newcommand{\apogee}{APOGEE}

\newcommand{\lims}{lower-intermediate mass stars}
\newcommand{\hms}{high-mass stars}

\defcitealias{james+21}{J21}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}

\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}

\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\texttt{VICE}}
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}

\newcommand{\alp}{$\alpha$}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
\newcommand{\Z}{Z}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}
\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\about}[1]{${\sim} #1$}

\newcommand{\proton}{\isotope[1][1]{p}}
\newcommand{\zoo}{\ensuremath{Z/Z_{\sun }}}
\newcommand{\Yct}{\ensuremath{y_{\rm C}}}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Origin and Galactic Evolution of C]{The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis }

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{
Daniel A. Boyea,$^{1}$\thanks{E-mail: boyea.2@osu.edu}
A. N. Other,$^{2}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Department of Astronomy, the Ohio State University, 191 W. Woodruff, Columbus, OH 43210, USA
}

@@

% Abstract of the paper
\begin{abstract}
% context
Limited knowledge of stellar physics leads to uncertainties in elemental yield predictions.
% aims
As an independent approach, we constrain the nucleosynthetic yields of C using multi-zone Galactic chemical evolution models.
% results
By matching the median trends of \caah\ and \caafe\ in  {\sc apogee} subgiant stars, we find that high-mass and low-mass stars contribute \about{80\%} and \about{20\%} of total C production respectively.
% 
To match the normalization of the trends, we estimate the massive star C/Mg yield ratio $\Ycc/\Yoc = 1.57 + 0.59 \left(Z/Z_\odot\right)$,  explaining the \caah\ trend when including the low mass C contribution.
%
Variations in the star formation history only slightly impact the low [Mg/Fe] tail of the [C/Mg]-[Mg/Fe] relation. However, most of the scatter in the \caah\ distribution can be attributed to measurement errors and radial migration. 
% 
Due to the degeneracy between the normalization of elemental yields and the strength of mass-loading in Galactic outflows, we am only able to constrain the relative yields of C/Mg and its metallicity dependence. While measurements of gas-phase C abundances are challenging, our model is broadly consistent with the [C/O]-[O/H] trend observed in compiled literature measurements.

\end{abstract}

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals\footnotemark{}. % \citep[e.g.][]{jennifer19, KL14}
C structurally changes the environments it pollutes, modifying stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% C/N as age indicator?
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

\footnotetext{Astronomers are not chemists. By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, $Z_\odot=0.014$. }

Galactic chemical evolution is a powerful tool in understanding the origin of elements. 
Each enrichment process has distinct chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation \citep{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20}.
(See also review in \citealt{romano22}.)
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metalliicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where carbon comes from and how it evolves, our aim is to use a combination of recent high-resolution observations from APOGEE and multi-zone models to develop estimates of the required C yields to reproduce galactic trends.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O are explained by the metallicity dependence of N/O yields. Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as \sfh{} and outflow mass loading.

We know, from observations of MW stars and extragalactic gas measurments, that C/O traces a banana shape in O/H (see \ref{fig:gas_phase}). 
% \cite[e.g.][]{fabbian+09, nissen+14, lambert81, laird85, lambert86}. 
At very low metallicities, C/O decreases with increasing metallicity, from observations of halo stars and damped Lyman-alpha systems\footnotemark{} \citep{nissen+14, cooke+17, fabbian+09}. 
\footnotetext{DLAs; high redshift clouds of gas observed as absorption lines in yet higher redshift quasars}
%
However, the trend reverses around $[{\rm O/H}]\approx-1$\footnotemark{}. 
C/O instead increases with increasing metallicity in observations of HII regions and Milky Way stars (\citealt{berg+19}; see discussion in section \ref{sec:gas}).

\footnotetext{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}

However, observations of C are not without challenges. 
As stars ascend the red giant branch (RGB) at the end of H core burning, C abundances are changed by \textit{first dredge up}\footnotemark{}. This has the benefit of allowing C/N to be used as an age indicator in RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}, but evolved stars are no longer as useful for determining birth abundances. Works such as \citet{vincenzo+21} attempt to correct for these changes but this requires MESA stellar modeling. 
In optical spectroscopy, C lines are relatively faint, and surveys such as GALAH struggle against low detection rates, potentially biasing sample measurments. While APOGEE includes more reliable C$_2$ lines, there is an additional risk of systematics because .....
In the gas phase, HII regions are our best window into C abundances in our galaxy and extragalactic regions including dwarf. 
Unfortunantly, C lacks strong optical emission lines. Approaches use either recombination lines or ... in the ultraviolet, but these methods disagree, making gas-phase abundances a limited metric for GCE models of C.

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17}. selected by the criteria in Roberts et al. (2023, in prep; see Appendix \ref{sec:jack}).
According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced first dredge up (FDU) but have well-mixed envelopes. So, subgiant surface abundances should represent their birth composition. 
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using these trends, we will develop a model of the enrichment sources and evolution of C.

\footnotetext{First dredge up (FDU) is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption[Subgiant Abundances]{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for the \citetjack~sample of \apogee{} subgiants. On the top, we plot high and low-$\alpha$ stars in blue and orange, using the separation defined in Equation \ref{eq:high_alpha} (the high and low-$\alpha$ stars are named for their high or low $\alpha$-element to Fe ratios, or in this case, Mg/Fe). On the bottom, we colour-code stars according to their [Mg/H] abundance.} \label{fig:subgiants}
\end{figure*}

\section{Nucleosynthesis}

Yields -- the amounts of each chemical element stars synthesise -- are central to studies of galactic chemical evolution (GCE). 
In this section, we discuss the yield choices of our model and literature models for C yields. 
We focus on three nucleosynthetic pathways: asymptotic giant branch (\agb{}) stars, core collapse supernovae (\cc{}), and type Ia supernovae (\ia{}).
C is produced in both \agb{} and \cc{} stars.
We also use Mg and Fe as tracers of \cc{} and \ia{} enrichment respectively. 
O and Mg are produced almost entirely from \cc\ with metallicity-independent yields, making each element. In contrast Fe is produced in similar amounts by \cc\ and \ia.

After a single stellar population (SSP)\footnotemark{} forms, \cc{} are the first chemical enrichers. \cc{} explode within $\lesssim 40$\,Myr, providing light elements (C, O, and Mg) and heavier elements (Fe and beyond). Next, low-mass stars begin to reach the end of their lives. By shedding their outer layers, \agb{} stars are important sources of C, N, and neutron capture elements.  Finally, white dwarfs explode in \ia{}, releasing Fe and other iron-peak elements.

\footnotetext{SSP: Single stellar population. Technical name for a group of stars born in the same conditions at the same time, i.e. an open cluster.}

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\tilde{y}$ is 
\begin{equation}
\tilde{y}_{X} = \frac{M_{\rm ejected}}{M_{\rm birth}} (Z_{f, X} -  Z_{0, X})
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $Z_{f, X}$ and $Z_{0, X}$ are the initial birth abundance and final abundance of the ejected envelop respectably. For example, if a 1\,\Mo\ star has $\tilde{y}_{\rm C} = 10^{-3}$, then the star will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. For a yield $y$ from a star of mass M and initial metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \tilde{y}_{\rm X}(M, Z)
    \frac{dN}{dM}  \ dM
\end{equation}
where ${dN}/{dM}$ is the normalized \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE\footnotemark{}).

We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law, as suggested by the observations of \citet{maoz+12}.

\addtocounter{footnote}{-2}
    \stepcounter{footnote}\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}}

\stepcounter{footnote}\footnotetext{\VICE~is available at \url{https://github.com/giganano/VICE}}

\begin{table}
	\centering
    \caption[Fiducial Model Yields]{Yields for the fiducial model (in units of \ssp~birth mass). See Section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l}
		\hline
        Element & $\Ycc$ & $\tilde{y}^{\rm AGB}$ & $y^{\rm SNe Ia}$ \\
		\hline
        C & $0.0028 + 0.001\,(Z/Z_\odot)$ & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & $9\times10^{-4}\,(Z/Z_\odot)\,M$ & 0\\
		\hline
	\end{tabular}
\end{table}

\subsection{Asymptotic Giant Branch Stars}\label{sec:agb}

An \agb\ star is a low-mass star ($\lesssim 8 M_{\sun}$) during its final phase of evolution.  
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \isotope[12]{C} into \isotope[14]{N} \citep{solar-fusion}.

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than about??2\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
In 2?? to 4?? \Mo stars, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In massive, high-metallicity \agb\ stars, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \isotope[12]{C} destruction.

\footnotetext{
    The CNO cycle and the details of AGB evolution are described in appendix \ref{sec:stel_evo}.
}

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following.
\begin{itemize}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{itemize}
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yield for these four models.
Note that the net yields may be negative if a star ejects material with a lower average C abundance than the material the star was formed from.
Most models agree on the qualitative shape of the net fractional \agb\ C yield---% 
stars with masses between about 2--4\,\Mo have the highest fractional C yields. As metallicity increases, the total AGB C yield decreases and the mass of the peak increases.

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\tilde{y}_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass AGB stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. In contrast, Fe production continues steadily for 10\,Gyr.

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity $Z$.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. All models predict yields within a factor of \about{2} for fixed metallicity.
For example, the three models \cxi, \kx{}, and \kxvi{} predict $y_\text{C}^\text{AGB}$ to be between 0.006 and 0.008 at solar metallicity, but \cxi\ has a much shallower metallicity dependence than the \kx{} and \kxvi{} models. \vxiii{} instead predicts a yield \about{0.004}.
    As shown in the right panel of Fig.~\ref{fig:agb-ssp}, with increasing $Z$, C enrichment occurs earlier, and C destruction in low-mass stars leads to a declining C abundance at late times.

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\tilde{y}=0$ for reference. Each panel represents yields from one of four \agb\ models: \cxi, \kx{}, \vxiii{}, \kxvi{} (see Section \ref{sec:agb} and  Appendix \ref{sec:oob_models}). }
        \label{fig:y_agb}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics{y_agb_vs_t.pdf}
    \includegraphics{y_agb_vs_z.pdf}

    \caption[]{ C production by \agb\ stars as a function of \ssp{} age, normalized to the total mass $M_{\rm tot}$ produced at $t=10$\,Gyr. \textbf{Left:} The four \agb\ yield models from literature at solar metallicity (\cxi, \kx{}, \vxiii{}, or \kxvi{}). The dashed black line shows the delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) for comparison. \textbf{Right:} The (\imf-weighted) \agb\ C yield $\Ycagb$ as a function of metallicity for each of the \agb\ yield models. ($\Ycagb$ is the net mass of C produced by \agb\ stars per unit mass of star formation, after 10\,Gyr and assuming a \citealt{kroupa01} \imf.) }

    \label{fig:agb-ssp}

@@

\subsection{Core Collapse Supernovae}

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process (see Appendix \ref{sec:stel_evo}). However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

Fig.~\ref{fig:y_cc} plots calculations of the \imf-integrated yields, defined with Eq.~\ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). 
\cc{} models predict a wide range of C yields, spanning almost a factor of ten. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
The \cite{LC18} models also include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.

Fig.~\ref{fig:y_cc} shows the \cxi{} \agb\ model for comparison on the left. Especially at $Z\approx Z_\odot$, most \cc{} models dominate \agb\ C production. Later, we will also show empirically this is the case.
On the right of Fig.~\ref{fig:y_cc}, we also show the \cc{} [C/Mg] ratio, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
If only \cc{} produced C, then ${\rm [C/Mg]^{CC}}$ describes the equilibrium abundance of [C/Mg].
Once again, different \cc\ models span a large range in [C/Mg]. 
we chose to instead parameterize $\Ycc$ to enable agreement with observations, as most \cc\ models fail to achieve near-solar [C/Mg].
Assumptions about the explodability landscape affect C and Mg production. Increasing the fraction of stars that explode increases $\Ycc$, as stars that directly collapse do not contribute to explosive yields \citep{emily+21}. However, C is relatively unaffected by the black-hole landscape, as very massive stars contribute C through enriched winds. Since Mg is formed deeper in the core of \hms, the Mg yield drops much more steeply, so models where few stars explode (S16/W18) have higher [C/Mg].
